Model_CFU

### 2016-12-16 Projects/HumanCdGF_cdiff_colonization.Rmd
adding Leave one cage out on all

@@

Is it possible one of the cages of the same donor are not similar, and causing the poor performance?

@@

By cage:  
`r summary(om_rsq_cage_df$rsq[om_rsq_cage_df$data_set=='test'])`

By donor:  
`r summary(om_rsq_pos_df$rsq[om_rsq_pos_df$data_set=='test'])`

>>>>

##### __Leave One Out__ - Model __Colonization level__ with Human Donor included
Model built using all OTUs present in >10% of samples, only displaying top 10% of predictive features

```{r LOCO}
cfu_all_src_shared_loo_df <- data.frame(colonized_cfu %>% 
    right_join(community_all_d0) %>% 
    filter(!is.na(log_cfu_med)) %>% 
    select(log_cfu_med, cage_id, mouse_id, one_of(names(top_features_src_cfu))))
source_list <- unique(cfu_all_src_shared_loo_df$cage_id)
cfu_all_src_shared_loo_df$human_source <- factor(cfu_all_src_shared_loo_df$human_source)

LOcO_probs <- c()
LOcO_rsq <- c()
for(i in 1:length(source_list)){
  by_source <- cfu_all_src_shared_loo_df$cage_id == source_list[i]
  LOcO_predict <- select(cfu_all_src_shared_loo_df, -mouse_id, -cage_id)
  train <- LOcO_predict[!by_source,]
  test <- LOcO_predict[by_source,]
  set.seed(seed)
  temp_model <- randomForest(log_cfu_med~., data=train, ntree=1001)
  LOcO_rsq[i] <- round(tail(temp_model$rsq, 1), 2)
  LOcO_probs <- rbind(LOcO_probs, data.frame(observed=test$log_cfu_med, 
                                             predicted=predict(temp_model, test),
                                             mouse_id=cfu_all_src_shared_loo_df$mouse_id[by_source],
                                             human_source=rep(source_list[i], nrow(test))))
}
LOcO_rsq_overall <- 1 - (sum((LOcO_probs$observed - LOcO_probs$predicted)^2)/
                   sum((LOcO_probs$observed - mean(LOcO_probs$observed))^2))
```

##### Leave one Cage out:  
Rsq of all predictions - `r LOcO_rsq_overall`  
Rsq of each iteration - `r LOcO_rsq`  
(cage order =  `r source_list`)

```{r LOSO}
cfu_all_src_shared_loo_df <- data.frame(colonized_cfu %>% 
    right_join(community_all_d0) %>% 
    filter(!is.na(log_cfu_med)) %>% 
    select(log_cfu_med, mouse_id, one_of(names(top_features_src_cfu))))
source_list <- unique(cfu_all_src_shared_loo_df$human_source)
cfu_all_src_shared_loo_df$human_source <- factor(cfu_all_src_shared_loo_df$human_source)

LOcO_probs <- c()
LOcO_rsq <- c()
for(i in 1:length(source_list)){
  by_source <- cfu_all_src_shared_loo_df$human_source == source_list[i]
  LOcO_predict <- select(cfu_all_src_shared_loo_df, -mouse_id)
  train <- LOcO_predict[!by_source,]
  test <- LOcO_predict[by_source,]
  set.seed(seed)
  temp_model <- randomForest(log_cfu_med~., data=train, ntree=1001)
  LOcO_rsq[i] <- round(tail(temp_model$rsq, 1), 2)
  LOcO_probs <- rbind(LOcO_probs, data.frame(observed=test$log_cfu_med, 
                                             predicted=predict(temp_model, test),
                                             mouse_id=cfu_all_src_shared_loo_df$mouse_id[by_source],
                                             human_source=rep(source_list[i], nrow(test))))
}
LOcO_rsq_overall <- 1 - (sum((LOcO_probs$observed - LOcO_probs$predicted)^2)/
                   sum((LOcO_probs$observed - mean(LOcO_probs$observed))^2))
```

#### Leave one Source out:  
Rsq of all predictions - `r LOcO_rsq_overall`  
Rsq of each iteration - `r LOcO_rsq`  
(source order =  `r source_list`)

Removing the repeated cages of DA00581 (A and C) causes the model to perform poorly leaving one out.
So is the performance of the whole model appears to be dependent on two grouping of points (one around ~5-6 log and one around ~8 log ). So how well can random forest model the data with out the DA00581 source?

```{r LOCO with out 581}
cfu_all_src_shared_loo_df <- data.frame(colonized_cfu %>% 
    right_join(community_all_d0) %>% 
    filter(!is.na(log_cfu_med)) %>% 
      filter(!human_source == 'DA00581') %>% 
    select(log_cfu_med, mouse_id, one_of(names(top_features_src_cfu))))
source_list <- unique(cfu_all_src_shared_loo_df$human_source)
cfu_all_src_shared_loo_df$human_source <- factor(cfu_all_src_shared_loo_df$human_source)

LOcO_probs <- c()
LOcO_rsq <- c()
for(i in 1:length(source_list)){
  by_source <- cfu_all_src_shared_loo_df$human_source == source_list[i]
  LOcO_predict <- select(cfu_all_src_shared_loo_df, -mouse_id)
  train <- LOcO_predict[!by_source,]
  test <- LOcO_predict[by_source,]
  set.seed(seed)
  temp_model <- randomForest(log_cfu_med~., data=train, ntree=1001)
  LOcO_rsq[i] <- round(tail(temp_model$rsq, 1), 2)
  LOcO_probs <- rbind(LOcO_probs, data.frame(observed=test$log_cfu_med, 
                                             predicted=predict(temp_model, test),
                                             mouse_id=cfu_all_src_shared_loo_df$mouse_id[by_source],
                                             human_source=rep(source_list[i], nrow(test))))
}
LOcO_rsq_overall <- 1 - (sum((LOcO_probs$observed - LOcO_probs$predicted)^2)/
                   sum((LOcO_probs$observed - mean(LOcO_probs$observed))^2))
```

#### Leaving one cage/source out (without Donor 581):  
Rsq of all predictions - `r LOcO_rsq_overall`  
Rsq of each iteration - `r LOcO_rsq`  
(cage/source order =  `r source_list`)

While this is removing 3 cages (8 mice) and decreasing the n, it is removing the smaller, lower grouping of points. So how well can the whole model predict CFU without the DA00581 samples?

```{r all without 581}
cfu_wo581_df <- colonized_cfu %>% 
    select(mouse_id, human_source, log_cfu_med) %>% 
    right_join(community_all_d0) %>% 
    filter(!human_source == 'DA00581') %>% 
    select(-mouse_id) %>% 
    filter(!is.na(log_cfu_med))
cfu_wo581_df$human_source <- factor(cfu_wo581_df$human_source)

set.seed(seed)
cfu_wo581_rf <- randomForest(cfu_wo581_df[ , -2], cfu_wo581_df$log_cfu_med,
                           ntree = 1001,
                           mtry = 15,
                           importance = T)
top_features_wo581 <- head(sort(importance(cfu_wo581_rf)[,1],decreasing = T), n_features)
```

```{r colonization performance without 581, fig.width=5, fig.height=4}
data.frame(Observed = cfu_wo581_df$log_cfu_med, 
           Predicted = cfu_wo581_rf$predicted) %>% 
  ggplot(aes(x = Predicted, y = Observed)) + geom_point() +
    xlim(5,9) + ylim(5,9) + 
    annotate('text', label = paste0(round(tail(cfu_wo581_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = min(cfu_wo581_rf$predicted), y = max(cfu_wo581_df$log_cfu_med)) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank()) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')

top_otus_no581 <- get_tax(row_list = names(top_features_wo581)[-1])$tax_label
top_otus_no581 <- c('Human Source', top_otus_no581)
```

```{r colonization otu importance without 581, fig.width=6, fig.height=6}
data.frame(OTU = factor(top_otus_no581, 
                        levels = top_otus_no581[order(top_features_wo581)]), 
           Importance = top_features_wo581/max(top_features_wo581)) %>% 
  ggplot(aes(x = OTU, y = Importance)) + 
    geom_point() + coord_flip() + 
    labs(y = 'Importance (Relative MDA)', x = NULL) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_line(colour = "gray", linetype = "dotted"),
          panel.grid.minor.y = element_line(colour = "gray", linetype = "dotted"),
          axis.text.y = element_text(color = c(rep('black', length(top_features_src_cfu)-13),
                                               rep('red', 12), 'black'), 
                                     hjust = 0))
```

Without human donor, performance was slightly lower with similar top features

__This decrease in performance when removing Human Donor 581 suggests that with this data set, random forest is unable to predict CFU with a high degree of accuracy. This model seems to potentially be identifying the split between the high and low CFU groups, although this is difficult to say with any weight since we only have one source in the low grouping.__

***
***

@@

Cages A, C, and D all have the same donor.  
Is it possible one of the cages of the same donor are not similar, and causing the poor performance?

@@

By cage:  
Min. 1st Qu.  Median    Mean 3rd Qu.    Max.   
`r summary(om_rsq_cage_df$rsq[om_rsq_cage_df$data_set=='test'])`

By donor:  
Min. 1st Qu.  Median    Mean 3rd Qu.    Max.

`r summary(om_rsq_pos_df$rsq[om_rsq_pos_df$data_set=='test'])`
